---
title: "New Template JC"
author: "John Corley"
date: "April 7, 2016"
output: html_document
---
# Loading and preprocessing the data
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)


### Load the data 
unzip("./activity.zip", "activity.csv")
file = read.csv("./activity.csv", header=TRUE)
str(file)

###Process/transform the data (if necessary) into a format suitable for your analysis
file$date <- as.Date(file$date)

```
# What is mean total number of steps taken per day?
```{r , echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)

### Calculate the total number of steps taken per day
dailySteps <- tapply(file$steps, file$date, sum, na.rm=TRUE)
str(dailySteps)
```
### Make a histogram of the total number of steps taken each day
``` {r, echo=TRUE}

#### Mean number of steps per day
myMean <- mean(dailySteps)
#### Median number of steps per day
myMedian <- median(dailySteps)

### Histogram, Median and Mean
hist(dailySteps, 25)
legend('topright', legend = c(paste("Mean =", round(myMean, 1)),paste("Median =",round(myMedian, 1))), bty = "n")

```

#What is the average daily activity pattern?
```{r , echo=TRUE}


###Time Series Plot

activity<-tapply(file$steps, file$interval, mean, na.rm=T)
plot(y = activity, x = names(activity), type = "l", xlab = "Nth 5 Minute Interval", main = "Daily Activity", ylab = "Avg. Num. Steps")

max_dim <- activity[which.max(activity)]
first <- round(as.numeric(levels(as.factor(max_dim))), 2)
second <- names(max_dim)
legend('topright', legend = c(paste("Time Interval: ", second),paste("Max Avg Steps: ",first)))
abline(v = as.numeric(names(max_dim)), col = 'red')
```

#Imputing missing values
####The total number of missing values is 
```{r , echo=TRUE}
sum(is.na(file$steps))
##which(is.na(file$steps))

```
Replacing 'NA' values with AVG Values per day, given the appropriate date of the 'NA'...
```{r , echo=TRUE}
copy <- file
copy$steps <- ifelse(is.na(copy$steps) == TRUE, as.numeric(levels(as.factor(activity)))[as.numeric(names(activity)) %in% copy$interval], copy$steps)
#df1$B <- ifelse(is.na(df1$B) == TRUE, df2$B[df2$A %in% df1$A], df1$B) 

```

##Histogram after Removed

```{r , echo=TRUE}
dailyStepsRm <- tapply(copy$steps, copy$date, sum, na.rm = TRUE)

#### Mean number of steps per day
myMeanRm <- mean(dailyStepsRm)
#### Median number of steps per day
myMedianRm <- median(dailyStepsRm)

hist(dailyStepsRm, 25)
legend('topright', legend = c(paste("Mean =", round(myMeanRm, 1)),paste("Median =",round(myMedianRm, 1))), bty = "n")
myMeanRm
```
# weekends vs weekdays
### you can see a skew for weekdays
```{r , echo =TRUE}

weekday <- copy[!(weekdays(copy$date) %in% c('Saturday', 'Sunday')),]
weekend <- copy[weekdays(copy$date) %in% c('Saturday', 'Sunday'),]

wd_activity <- tapply(weekday$steps, weekday$interval, mean)
we_activity <- tapply(weekend$steps, weekend$interval, mean)
par(mfrow=c(2,1))
plot(wd_activity, type="l", main="Weekdays", xlab="Nth 5min Interval", ylab="Avg Steps")
plot(we_activity, type="l", main="Weekends", xlab="Nth 5min Interval", ylab="Avg Steps")
```


